---
title: "p8105_hw3_yl4613 codes"
author: "YuechenLiu"
date: "10/9/2020"
output: html_document
---


```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
	fig.width = 6,
	fig.asp = .6,
	out.width = '90%'
)

theme_set(theme_minimal() + theme(legend.position = 'bottom'))

options(
	ggplot2.continuous.colour = 'viridis',
	ggplot2.continuous.fill = 'viridis'
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` and `r ncol(instacart)` columns.

Observations are the level of items in order by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes.

How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```


Let's make a plot

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>%
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) +
	geom_point() +
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


Let's make a table!!

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


Apples vs ice cream..

```{r}
instacart %>% 
	filter(product_name %in% c('Pink Lady Apples', 'Coffee Ice Cream')) %>%
	group_by(product_name, order_dow) %>%
	summarize(mean_hour = mean(order_hour_of_day)) %>%
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)

```


### Problem 2
### load and tidy the data
```{r}
ac_data = read.csv(
	'./accel_data.csv') %>% 
	janitor::clean_names() %>% 
	pivot_longer(activity_1:activity_1440,
							 names_to = 'minute_of_the_day',
							 values_to = 'activity_counts',
							 names_prefix = 'activity_') %>% 
	mutate(
		type_of_day = case_when(
			day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday') ~ 'weekday',
			day %in% c('Saturday', 'Sunday') ~ 'weekend'),
			day = as.factor(day),
		minute_of_the_day = as.integer(minute_of_the_day), 
		type_of_day = as.factor(type_of_day))
```
The resulting data is tidier than the original dataset. It contains: number of the week (1-5), day_id(useless now so I left it alone), name of the data (Monday-Sunday), minute of the day (1-1440, every minute of a single day), activity_counts(the activity counts for each minute of that day), type_of_day(Monday-Friday are defined as weekdays, Saturday and Sunday are defined as weekends). There are `r nrow(ac_data)` observations in total, which contains 5 weeks, 35 days. So every day there are `r nrow(ac_data)/35` observations.

Traditional analyses--total activity of the day.
```{r}
sum_data = 
	tibble(ac_data) %>% 
	group_by(day, week) %>% 
	summarize(activity_of_day = sum(activity_counts)) %>% 
	relocate(week) %>% 
	arrange(week) %>% 
	pivot_wider(
		names_from = day,
		values_from = activity_of_day) %>% 
	relocate(week, Sunday, Monday, Tuesday, Wednesday, Thursday, Friday, Saturday)
sum_data
```
Trends: The activity data during the middle of the week is more stable than that in weekends. For every week, week 1-3 have the most activities, whereas activities reduced during the last two weeks. 

Plotting-24 hour activity
```{r}
ac_data %>% 
	ggplot(aes(x = minute_of_the_day, y = activity_counts , color = day)) +
	geom_line(alpha = .4) + 
	labs(x = 'minute of the day',
			 y = 'numbers of activity',
			 title = 'Activities every minute during a day')
```
From this plot, we can see that the activities during daytime are higher, and during the night are lower. It indicates that this patient has a daily routine that: wakes up ~7-8 am, have some activities before noon. He might take a nap / have some rest in the afternoon. He has high activity during ~8-11pm. And he sleeps usually after 11pm.


###Problem 3
```{r}
library(dplyr)
library(rnoaa)
```





























